Accelerated Bilateral Filtering for Many-core and Multi-core Architectures

### 2012-03-11 trunk/LateX/paper8.tex
changed manycore to many-core.

\begin{document}
\title{Acceleration of Bilateral filtering algorithm for manycore and multicore architectures}
\author{\IEEEauthorblockN{Dinesh Agarwal, Sami Wilf, Abinashi Dhungel, and Sushil K. Prasad}\\
\IEEEauthorblockA{ Georgia State University\\
 Department of Computer Science\\
 Atlanta, Georgia 30303\\
 }}
\maketitle
\begin{abstract}
This work explores multicore and manycore acceleration for the embarrassingly parallel, compute-intensive Bilateral filtering kernel. For manycore architectures, we have created a pair-symmetric algorithm to avoid redundant calculations and for multicore architectures we improve the algorithm by use of low level single instruction multiple data (SIMD) parallelism across multiple threads. We present empirical data evidencing the performance gains we achieved over variety of hardware architectures including Nvidia GTX280, AMD Barcelona, AMD Shanghai, Intel Harpertown, AMD Phenom, Intel Core i7 quad core, and Intel Nehalem 32 core machines. The best speedup achieved was a 235.5x speedup by our CUDA-based implementation of our pair-symmetric algorithm running on Nvidia's GTX280. Our CPU multicore implementations resulted in a speedup of up to 38x using 16 cores of AMD Barcelona each with 4-stage vector pipeline, up to 23x using 8 cores for Intel Harpertown, and up to 26x using 32 cores of Intel Nehalem multicore machines compared to a compiler-optimized code.
\end{abstract}
% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle
\section{Introduction}
\label{sec:intro} The recent advances in parallel computing and architecture design have provided a series of avenues for application performance improvements. However, unlike the days when applications automatically became faster with increased clock frequency, newer chips require programmers and researchers to revisit their algorithm designs to harness the power of underlying architectures. Moreover, there is no well defined list of guidelines that can be followed for each application, which makes it challenging to deploy existing applications, even fundamental ones, to a myriad of available multicore and manycore architectures.

@@

\subsection{Shared Memory and Tile Processing}
\begin{figure}
\includegraphics[width = 0.48 \columnwidth]{images/project24.pdf}
\caption{Illustration of our CUDA implementation of pair-symmetric processing using implicit intra-wap synchronization for iterating through the pixel-cells for each thread-block's assigned tile.}
\label{fig:pspro}
\end{figure}

\begin{figure}
\includegraphics[width = 0.48  \columnwidth]{images/tiles19.pdf}
\caption{Illustration of our CUDA implemented pair-symmetric algorithm processing a 24 x 12 pixel image using 2 x 2 thread-blocks each with 4 x 3 threads and with the half-width parameter set to 3 pixels.}
\label{fig:pspro}
\end{figure}

\begin{figure}
\includegraphics[width = 0.48  \columnwidth]{images/smallfaces5.pdf}
\caption{An illustrated overview of the bilateral filtering process. The six raster images of processed tiles reflect the successive work of the four concurrent thread-blocks iteratively carried out over the padded source-image.}
\label{fig:pspro}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.48\textwidth]{images/stackedtiles}
\caption{Tiling for Pair-Symmetric Stencil Processing with each tile having a unique color and overlapping tiles have a resultant color that results from mixing colors of overlapping tiles. The pair-symmetric thread access pattern ensures exclusive memory writes.}
\label{fig:Tiling-for-Pair-Symmetric}
\end{figure}

\begin{figure}
\includegraphics[width=0.48\textwidth]{images/lenatiles}
\caption{Input image after one iteration of pair-symmetric algorithm. Tiles do not have perfect boundaries due to overlaps.}
\label{fig:An-iteration-of}
\end{figure}

@@

Another challenge in interpreting our results was porting the concept of "speed-up" to general purpose GPU programming. Since speed-up is necessarily with respect to a reference timing, we used the same reference timing that our CPU optimizations used---that is how we calculated the 235.5x speed-up figure stated in the abstract of this paper---and we also made another reference timing by running our GPU implementations using a single thread. For example, using a single thread operating with filter width of 10 pixels to process a 768 x 512 pixel image, the time taken for the GTX280 to complete this execution was an average of 177.9 seconds for both of our algorithms; using a filter width of 5 pixels, the average execution time was 50.5 seconds for both algorithms. With respect to these reference timings, the speed-up is roughly 20,500x for the 10 pixel filter width and 12,440x for the 5 pixel filter width.

>>>>

\begin{document}
\title{Accelerated Bilateral Filtering for Many-core and Multi-core Architectures}
\author{\IEEEauthorblockN{Dinesh Agarwal, Sami Wilf, Abinashi Dhungel, and Sushil K. Prasad}\\
\IEEEauthorblockA{ Georgia State University\\
 Department of Computer Science\\
 Atlanta, Georgia 30303\\
 }}
\maketitle
\begin{abstract}
This work explores multicore and many-core acceleration for the embarrassingly parallel, compute-intensive Bilateral filtering kernel. For many-core architectures, we have created a pair-symmetric algorithm to avoid redundant calculations and for multicore architectures we improve the algorithm by use of low level single instruction multiple data (SIMD) parallelism across multiple threads. We present empirical data evidencing the performance gains we achieved over variety of hardware architectures including Nvidia GTX280, AMD Barcelona, AMD Shanghai, Intel Harpertown, AMD Phenom, Intel Core i7 quad core, and Intel Nehalem 32 core machines. The best speedup achieved was a 235.5x speedup by our CUDA-based implementation of our pair-symmetric algorithm running on Nvidia's GTX280. Our CPU multicore implementations resulted in a speedup of up to 38x using 16 cores of AMD Barcelona each with 4-stage vector pipeline, up to 23x using 8 cores for Intel Harpertown, and up to 26x using 32 cores of Intel Nehalem multicore machines compared to a compiler-optimized code.
\end{abstract}
% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle
\section{Introduction}
\label{sec:intro} The recent advances in parallel computing and architecture design have provided a series of avenues for application performance improvements. However, unlike the days when applications automatically became faster with increased clock frequency, newer chips require programmers and researchers to revisit their algorithm designs to harness the power of underlying architectures. Moreover, there is no well defined list of guidelines that can be followed for each application, which makes it challenging to deploy existing applications, even fundamental ones, to a myriad of available multicore and manycore architectures.

@@

\subsection{Shared Memory and Tile Processing}
\begin{figure}
\includegraphics[width = 1\columnwidth]{images/project23_.pdf}
\caption{Illustration of our CUDA implementation of pair-symmetric processing using implicit intra-wap synchronization for iterating through the pixel-cells for each thread-block's assigned tile.}
\label{fig:pspro}
\end{figure}

\begin{figure}
\includegraphics[width = 1\columnwidth]{images/tiles19.pdf}
\caption{Illustration of our CUDA implemented pair-symmetric algorithm processing a 24 x 12 pixel image using 2 x 2 thread-blocks each with 4 x 3 threads and with the half-width parameter set to 3 pixels.}
\label{fig:pspro}
\end{figure}

\begin{figure}
\includegraphics[width = 1\columnwidth]{images/smallfaces5.pdf}
\caption{An illustrated overview of the bilateral filtering process. The six raster images of processed tiles reflect the successive work of the four concurrent thread-blocks iteratively carried out over the padded source-image.}
\label{fig:pspro}
\end{figure}

@@

%\begin{figure}
%\includegraphics[width=0.48\textwidth]{images/stackedtiles}
%\caption{Tiling for Pair-Symmetric Stencil Processing with each tile having a unique color and overlapping tiles have a resultant color that results from mixing colors of overlapping tiles. The pair-symmetric thread access pattern ensures exclusive memory writes.}
%\label{fig:Tiling-for-Pair-Symmetric}
%\end{figure}

%\begin{figure}
%\includegraphics[width=0.48\textwidth]{images/lenatiles}
%\caption{Input image after one iteration of pair-symmetric algorithm. Tiles do not have perfect boundaries due to overlaps.}
%\label{fig:An-iteration-of}
%\end{figure}

@@

Another challenge in interpreting our results was porting the concept of "speed-up" to general purpose GPU programming. Since speed-up is necessarily with respect to a reference timing, we used the same reference timing that our CPU optimizations used---that is how we calculated the 235.5x speed-up figure stated in the abstract of this paper---and we also made another reference timing by running our GPU implementations using a single thread. For example, using a single thread operating with filter width of 10 pixels to process a 768 x 512 pixel image, the time taken for the GTX280 to complete this execution was an average of 177.9 seconds for both of our algorithms; using a filter width of 5 pixels, the average execution time was 50.5 seconds for both algorithms. With respect to these reference timings, the speed-up is roughly 21,090x for the 10 pixel filter width and 12,440x for the 5 pixel filter width.
